---
title: "PA1_template.Rmd"
author: "Jason Newbury"
date: "November 14, 2015"
output: html_document
---
PA1_Template
================================================================

First we must read in the data

```{r}
activity <- read.csv("activity.csv")

```

Graph a histogram of the data
```{r}
library(plyr)
day_activity <- ddply(activity,.(date),function(x) sum(x$steps))
colnames(day_activity)[2] <- "steps"
hist(day_activity$steps, main = "Histogram of Total Steps per Day" , xlab = "Number of Steps")
```
  
  Find the mean
```{r}
mean(day_activity$steps)
```
  Find the median
```{r}
median(day_activity$steps)
```

If we strip the NAs we get the following for the mean
```{r}
mean(day_activity$steps,na.rm = TRUE)
```
and the median
```{r}
median(day_activity$steps,na.rm = TRUE)
```

Now we find out the average number of steps taken in each 5 minute period
```{r}
min_activity <- ddply(activity, .(interval), function(x) mean(x$steps, na.rm=TRUE))
colnames(min_activity)[2] <- "steps"
plot(min_activity$interval,min_activity$steps, xlab = "Minutes",main = "Average Number of Steps During Each 5 Minute Interval", ylab = "Average Number of Steps", type = "l")
```

Find at what time the max number of steps occur
```{r}
min_activity$interval[which.max(min_activity$steps)]
```


Find the number of NA's
```{r}
sum(is.na(activity$steps))
```

Replace any NA with the average value for that 5 min interval
```{r}
nas<- is.na(activity$steps)
average_interval <- tapply(activity$steps, activity$interval, mean,na.rm=TRUE)
activity$steps[nas] <- average_interval[as.character(activity$interval[nas])] 
```

Redo the histogram from earlier with the NA less data
```{r}
day_activity <- ddply(activity,.(date),function(x) sum(x$steps))
colnames(day_activity)[2] <- "steps"
hist(day_activity$steps, main = "Histogram of Total Steps per Day" , xlab = "Number of Steps")
mean(day_activity$steps)
median(day_activity$steps)
```

Find add a column for weekend or weekdays
```{r fig.height=10}
activity <- mutate(activity, weekdayorweekend = ifelse(weekdays(as.Date(activity$date))== "Saturday" | weekdays(as.Date(activity$date)) == "sunday", "Weekend", "Weekday" ))
par(mfrow= c(2,1))
activity_weekday <- subset(activity, weekdayorweekend == "Weekday")
activity_weekend <- subset(activity, weekdayorweekend == "Weekend")

min_activity_weekend <- ddply(activity_weekend, .(interval), function(x) mean(x$steps, na.rm=TRUE))
colnames(min_activity_weekend)[2] <- "steps"
plot(min_activity_weekend$interval,min_activity_weekend$steps, xlab = "Minutes",main = "Weekend", ylab = "Average Number of Steps", type = "l")

min_activity_weekday <- ddply(activity_weekday, .(interval), function(x) mean(x$steps, na.rm=TRUE))
colnames(min_activity_weekday)[2] <- "steps"
plot(min_activity_weekday$interval,min_activity_weekday$steps, xlab = "Minutes",main = "Weekday", ylab = "Average Number of Steps", type = "l")
```

Seems people get start moving earlier on Weekdays and stop moving sooner.

